---
output: github_document
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/",
  dpi=300
)
```

# IndexNumberTools

**IndexNumberTools** intends to ease the everyday work of users and producers of index numbers by providing functionalities like chain-linking, base shifting or computing pyp indices.

## Installation

```{r, eval = FALSE}
# Install released version from CRAN
install.packages("IndexNumberTools")
```

## Getting started

We'll work through an example using the [Spanish GDP](https://ine.es/dynt3/inebase/en/index.htm?padre=11371&capsel=11371) in current prices, which is preloaded as `gdp_current`, and the chain-linked volume (index), `gdp_volume`. 

```{r include=FALSE}
library(IndexNumberTools)
```

We can easily re-reference the volume series with `change_ref_year()`.[^slice]

[^slice]: We show only a slice of the output.

```{r}
gdp_volume_2010 <- change_ref_year(gdp_volume, 2010)
```

```{r, echo=FALSE}
volume <- cbind(aggregate.ts(gdp_volume, FUN = mean),
                aggregate.ts(gdp_volume_2010, FUN = mean))
colnames(volume) <- c("2020", "2010")
window(volume, start = 2008, end = 2022)
```

We can also get the series at previous year prices from `gdp_volume` with  `get_pyp()`.[^annual]

[^annual]: See the article "Annual arguments" for a detailed discussion.

```{r}
gdp_pyp <- get_pyp(gdp_volume)
```

```{r, echo = FALSE}
plot(gdp_pyp, main = "GDP pyp index", ylab = "index")
```

Multiplying the volume series by the mean of the current prices series at the reference year (2020), we obtain the GDP in (chain-linked) constant prices.

```{r}
ref_year_mean <- window(gdp_current,start = c(2020,1), end = c(2020,4)) |> mean()
gdp_constant <- ref_year_mean * gdp_volume / 100
```

By dividing the GDP in current prices by the GDP in constant prices, we derive the chain-linked implicit deflator of the GDP.

```{r}
gdp_deflator <- gdp_current / gdp_constant * 100
```


Using `get_v_index()` and chain-linking the result with `get_chain_linked()`, we get the chain-linked value indices.

```{r}
gdp_value <- get_v_index(gdp_current) |> get_chain_linked(2020)
```

Then, we can verify the identity $V = P\cdot Q$, that is, the value index must equal the product of the price and volume indices.

```{r message=FALSE, warning=FALSE}
dplyr::near(gdp_value, gdp_deflator * gdp_volume / 100) |> all()
```

```{r echo=FALSE}
plot(gdp_volume, ylab = "index", col = "green", main = "Volume, price and value indices")
lines(gdp_deflator, col = "red")
lines(gdp_value, col = "blue")
lines(gdp_deflator * gdp_volume / 100, col= "orange", lty=2)
legend(x = "bottomright",      # Posición
       legend = c("volume", "price", "value", "volume*price"), 
       col = c("green", "red", "blue", "orange"),
       lty = c(1,1,1,2))
```


