---
title: "1. Overview with Spanish GDP"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{1. Overview with Spanish GDP}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(IndexNumberTools)
```


The [Spanish GDP](https://ine.es/dynt3/inebase/en/index.htm?padre=11371&capsel=11371) in current prices is preloaded with the package as `gdp_current`, and the chain-linked volume (index) as `gdp_volume`. Multiplying the volume series by the mean of the current prices series at the reference year (2020), we obtain the GDP in constant prices.


```{r}
ref_year_mean <- window(gdp_current,start = c(2020,1), end = c(2020,4)) |> mean()
gdp_constant <- ref_year_mean * gdp_volume / 100
```

```{r echo=FALSE}
plot(gdp_current, ylab="GDP")
lines(gdp_constant, col = "red")
legend(x = "bottomright",      # Posición
       legend = c("current prices", "constant prices"), 
       col = c("black", "red"),
       lty = c(1,1))
```

By dividing the GDP in current prices by the GDP in constant prices, we derive the chain-linked implicit deflator of the GDP. Using `get_v_index()` and chain-linking the result with `get_chain_linked()`, we get the chain-linked value indices.

```{r}
gdp_deflator <- gdp_current / gdp_constant * 100
gdp_value <- get_v_index(gdp_current) |> get_chain_linked(2020)
```

Then, we can verify the identity $V = P\cdot Q$, that is, the value index must equal the product of the price and volume indices.

```{r message=FALSE, warning=FALSE}
dplyr::near(gdp_value, gdp_deflator * gdp_volume / 100) |> all()
```

```{r echo=FALSE}
plot(gdp_volume, ylab = "index", col = "blue")
lines(gdp_deflator, col = "red")
lines(gdp_value, col = "green")
legend(x = "bottomright",      # Posición
       legend = c("volume", "price", "value"), 
       col = c("blue", "red", "green"),
       lty = c(1,1,1))
```
